ProSolo: Accurate Variant Calling from Single Cell DNA Sequencing Data

### 2021-02-02 manuscript/main.tex
last updates to text, e.g. citing the general Zenodo DOI that points to latest pipeline and results data version

\newcommand{\Prob}{{\mathbf{P}}}
\newcommand{\cR}{{\mathfrak R}}
\newcommand{\cB}{{\mathbf{B}}}

\usepackage{euscript}

@@

Thus, variant callers for whole-genome-amplified single-cell data need to account for both amplification errors and allelic bias, in addition to accounting for the site-specific variation.
However, state of the art single cell SNV callers routinely assume fixed global rates when modeling uneven allelic coverage (up to dropout) and amplification errors. 
For example, to reflect the amplification error rate, both MonoVar\cite{zafar_monovar:_2016} and SCcaller\cite{dong_accurate_2017} work with global false positive error rates for calling the presence of an alternative allele at a particular site.
This assumes that $\Phi29$ polymerase is agnostic to local template sequence context, although it is not\cite{de_paz_high-resolution_2018}.
Similarly, for modelling allele dropout, MonoVar\cite{zafar_monovar:_2016} and SCIPhI\cite{singer_single-cell_2018} assume that one rate applies globally (and across all cells).
This neglects that allele dropout, as the extreme case of uneven allele coverage, varies greatly along the genome and in particular also between cells, because their DNA is amplified separately.
Interestingly, SCIPhI\cite{singer_single-cell_2018} additionally models allelic amplification bias to be governed by one global beta-binomial distribution (to apply for all cells in a dataset), thereby accounting for allelic dropout a second time (as the extreme values at 0 and 1 of that distribution).
Two tools that more variably model alleleic amplification bias are SCcaller \cite{dong_accurate_2017} and SCAN-SNV\cite{luquette_identification_2019}.
Both estimate the minor allele frequency from nearby germline heterozygous sites. 
However, SCcaller employs a fixed global false positive error rate for the calling of alternative alleles\cite{dong_accurate_2017}, and SCAN-SNV makes use of heuristics for filtering candidate variants\cite{luquette_identification_2019}.
Thus, to the best of our knowledge, there is no statistical model that allows for local variation of bias and errors due to amplification, and for statistically sound false discovery control when calling and genotyping SNVs in single cells.

@@

@@

\paragraph{Calculating posterior probabilities for events at single cell sites, including a bulk background sample.}
With single cell genotype likelihoods adjusted by our empirical amplification bias model (Equation \ref{eq:BBLodato}) and auxiliary evidence on alternative alleles from a bulk sample (Equation \ref{eq:bulk-prob}), we define mutually exclusive single cell events.
Figure~\ref{fig:prosolo_alt-calling}C gives a simplified illustration of how the combination of likelihood density estimates across allele frequency $\tilde{\theta}_s,\tilde{theta}_b$ works for calling a heterozygous and a homozygous genotype in the single cell.
However, our model fully defines the two-dimensional space of possible underlying alternative allele frequencies in the two samples as:

@@

We compare ProSolo to the state-of-the-art for SNV calling from single cell sequencing data of multiple displacement amplified (MDA) DNA: MonoVar\cite{zafar_monovar:_2016}, SCAN-SNV\cite{luquette_identification_2019}, SCcaller\cite{dong_accurate_2017} and SCIPhI\cite{singer_single-cell_2018}. We used Snakemake\cite{koster_snakemakescalable_2012} (version {\ttfamily 5.4.0}) to implement the benchmarking workflows.
For detailed information on benchmarking setup and results, see Supplementary Section~\ref{sec:suppl-benchmarking}.
All code used for benchmarking is available at: \url{https://github.com/prosolo/benchmarking_prosolo} (or as preserved by Zenodo at: \url{https://doi.org/10.5281/zenodo.3824152}).

@@

\paragraph{Whole exome sequencing of five human granulocytes with a pedigree ground truth.}
For the second benchmarking dataset, blood was taken from a patient with a constitutional mismatch repair-deficiency\cite{hoell_constitutional_2014} after informed consent.
Granulocytes were selected via Magnetic-Activated Cell Sorting (MACS) and Fluorescence Activated Cell Sorting (FACS, Figure~\ref{fig:datasets}B, dataset available as \mbox{EGAD00001005929} in EGA study \mbox{EGAS00001004123}).
Individual cells were isolated using a microfluidics device and subjected to multiple displacement amplification (MDA).
Using a panel of 16 primer pairs covering different genes across chromosomes for quantitative real-time PCR, we selected granulocytes where at least 15 of these loci were properly amplified.
For those cells, we performed whole exome capture, sequencing library preparation and paired-end Illumina sequencing.
From the remaining sorted cell population, we also extracted bulk DNA and submitted it to whole exome capture and paired-end Illumina sequencing without MDA, to generate a bulk background sample for ProSolo and SCcaller.

@@

\paragraph{Single nucleus exome sequencing of 16 normal and 16 tumor cells from a triple negative breast cancer (TNBC)\cite{wang_clonal_2014}.}
The third dataset is a published dataset from a TNBC patient, with both tumor and matched normal cells available (Figure~\ref{fig:datasets}C, dataset available via the sequence read archive, project id \mbox{PRJNA168068} or accession \mbox{SRA053195}).
For each, 16 single cell nuclei and a bulk sample were whole exome sequenced.
Further, a number of candidate sites for clonal and subclonal tumor mutations was confirmed by targeted duplex deep sequencing in a separate experiment\cite{wang_clonal_2014}.
We analyzed the tumor and normal cells separately, ensuring that normal cells were called with the normal bulk background sample and tumor cells with the tumor bulk background sample.
As the ground truth for the tumor cells, we then used the normal bulk sample augmented with the clonal tumor mutations confirmed by targeted duplex sequencing.
As the ground truth for the normal cells, we used the tumor bulk sample, removing the confirmed clonal and subclonal tumor mutations. Note that this experimental setup accounts for utmost fairness for all competitors. 
Variant calling for the ground truths was performed as for cell line data above.

@@

 \caption{
  Precision-recall plots for alternative allele calls of ProSolo, MonoVar, SCAN-SNV, SCcaller and SCIPhI.
  Note that all panels are strong zoom-ins, focusing on (different) areas of interest.
  Global views of these panels are provided in Figure~\ref{fig:alt-calling_prec-rec_global}.
  \textbf{A} Precision and recall average of two whole genome sequenced single cells IL-11 and IL-12 against their kindred clone IL-1C as ground truth genotypes.
  \textbf{B} Precision and recall average of the five whole exome sequenced single granulocytes against their pedigree-based germline genotype ground truth.
  \textbf{C} Precision and recall average of 16 tumor and 16 normal single cells sequenced at the whole exome level. \newline
  $^\text{\sffamily-B}$ The germline ground truth induces an artificial increase of recall for SCIPhI's sensitive and ProSolo's imputation mode; these modes should thus be disregarded for a fair comparison on the granulocyte dataset in panel B. \newline \footnotesize
  Threshold parameters:
  MonoVar {\ttfamily --t};
  ProSolo {\ttfamily --fdr};
  SCAN-SNV {\ttfamily --fdr};
  SCcaller {\ttfamily -a cutoff};
  SCIPhI {\ttfamily prosolo --fdr}.
  Software modes:
  MonoVar with consensus filtering ({\itshape default}) or without ({\itshape no consensus});
  ProSolo with minimum coverage 1 in single cell ({\itshape default}), or imputing zero coverage sites based on bulk sample ({\itshape imputation});
  SCcaller with recommended settings ({\itshape default}) or with a more {\itshape sensitive} calling;
  SCIPhI with default parameters ({\itshape default}) or all heuristics off ({\itshape sensitive}).
 }
\label{fig:alt-calling_prec-rec}
\end{figure}

@@

At the same time, this germline ground truth caveat also indicates that the recall of SCIPhI's sensitive mode and ProSolo's imputation mode will be an overestimate.
Whenever coverage of a site is missing in a single cell, SCIPhI may impute the genotype with the last common ancestor genotype of the most closely related cells, while ProSolo will impute to the majority genotype in the bulk sample.
Both strategies provide a biologically meaningful imputation that will be more useful than post-hoc modes of imputation.
However, at single cell sites where a somatic mutation has created a true alternative allele, but no coverage is provided, we expect that both methods are most likely to call the homozygous reference germline genotype.
In then comparing this to the germline genotype as the ground truth, these calls will be classified as true negatives even though they really constitute false negatives, thus artificially increasing recall.
While the underestimation of precision equally affects all tools and generally means that benchmarking results are more conservative than with a more accurate (somatic) ground truth, this overestimation of recall in only these modes of two tools does not allow for a fair comparison.
We have thus excluded both SCIPhI's sensitive mode and ProSolo's imputation mode from the discussion of the whole exome dataset with its germline ground truth (but their results are nevertheless displayed in Figures~\ref{fig:alt-calling_prec-rec}B, \ref{fig:alt-calling_prec-rec_global}B, \ref{fig:F_1-scores}B and \ref{fig:FDR-ground-truth-vs-theoretical}B for reference).

For the third dataset of single nucleus whole exome sequencing data of 32 cells from a TNBC patient (16 tumor cells, 16 normal cells), we analysed tumor and normal cells separately (Figure~\ref{fig:alt-calling_prec-rec}C).
On the tumor cells, ProSolo is the only tool to achieve a precision above .99 (at a recall of .319).
On the normal cells, both ProSolo and SCIPhI achieve a precision above .99.
Here, SCIPhI outcompetes ProSolo with a maximum recall of .650 compared to ProSolo's .548.
While ProSolo can achieve a similar recall (.625, see {\ttfamily --fdr} threshold of 0.2 in Figure~\ref{fig:alt-calling_prec-rec_global}C), this comes at a drastically reduced precision.
This reflects the inherent uncertainty of the single cell data and showcases a key difference in the approaches of ProSolo and SCIPhI.
Where ProSolo models each cell and each genomic site separately, SCIPhI's model integrates information across all sites in all cells at once.
While this can clearly help recover recall, this can also lead to false positive somatic variant calls (Figure~\ref{fig:somatic-recall}) and has clear implications for model complexity and software runtime.
As all other tools, including ProSolo, are parallelizable over genomic regions and were thus able to process both datasets within days on a multicore machine, we do not report more detailed runtimes for them.
However, it should be noted that SCIPhI took from 1 week (with iterations reduced below software defaults) on the single nucleus whole exome dataset up to 7.5 weeks on the whole genome dataset, running on a single core without any possibility of parallelization.
And where adding breadth of coverage (i.e. more genomic sites) or more cells simply means adding more parallel processes in most tools, both will further increase SCIPhI's wall time.
In addition, adding more cells exponentially grows the space of possible tree topologies that SCIPhI explores\cite{lahnemann_eleven_2020}, which will further increase its runtime.

@@

ProSolo is the first method for SNV calling from MDA single cell sequencing data to comprehensively model both amplification bias and amplification errors in a way that enables both
\begin{enumerate*}
 \item the integration of site-specific differences (Figure \ref{fig:prosolo_alt-calling}), and
 \item a computationally efficient---hence practically scalable---evaluation of all relevant data properties.
\end{enumerate*}
We achieve this by combining a data-driven model of amplification errors (that incorporates a bulk background sample alongside the single cell sample) with an empirical model of amplification bias which is dependent upon a site's coverage (and is based on a statistical understanding of the MDA process that is mechanistically motivated).
The underlying model calculates posterior probabilities for fine-grained single cell event definitions, whose false discovery rate can be controlled for and that can pass more information about data uncertainties to probabilistic models in downstream analyses; including methods that can then use this reduced, but highly accurate, data representation to integrate information across cells, for example for the phylogenetic reconstruction of cell relationships.
Importantly, the calculation of posterior probabilities scales linearly in the coverage of variant sites, which is the fastest theoretically conceivable, documented by runtimes that have substantial advantages over the state of the art in practice.
Using one whole genome dataset and two whole exome datasets---each with a different type of ground truth (Figure~\ref{fig:datasets})---we demonstrate that these two innovations of ProSolo combined to achieve competing or better accuracy (Figure~\ref{fig:alt-calling_prec-rec}).
Moreover, the model allows to accurately control the false discovery rate of variant calls, another novelty of great practical value, and both the model and its modular implementation are easy to adapt and extend.

>>>>

\newcommand{\Prob}{{\mathbf{P}}}
\newcommand{\cR}{{\mathfrak R}}
\newcommand{\cB}{{\mathbf{B}}}
\usepackage{euscript}

@@

Thus, variant callers for whole-genome-amplified single-cell data need to account for both amplification errors and allelic bias, in addition to accounting for the site-specific variation.
However, state of the art single cell SNV callers routinely assume fixed global rates when modeling uneven allelic coverage (up to dropout) and amplification errors. 
For example, to reflect the amplification error rate, both MonoVar\cite{zafar_monovar:_2016} and SCcaller\cite{dong_accurate_2017} work with global false positive error rates for calling the presence of an alternative allele at a particular site.
This assumes that $\Phi29$ polymerase is agnostic to local template sequence context, although it is not\cite{de_paz_high-resolution_2018}.
Similarly, for modelling allele dropout, MonoVar\cite{zafar_monovar:_2016} and SCIPhI\cite{singer_single-cell_2018} assume that one rate applies globally (and across all cells).
This neglects that allele dropout, as the extreme case of uneven allele coverage, varies greatly along the genome and in particular also between cells, because their DNA is amplified separately.
Interestingly, SCIPhI\cite{singer_single-cell_2018} additionally models allelic amplification bias to be governed by one global beta-binomial distribution (to apply for all cells in a dataset), thereby accounting for allelic dropout a second time (as the extreme values at 0 and 1 of that distribution).
Two tools that more variably model allelic amplification bias are SCcaller \cite{dong_accurate_2017} and SCAN-SNV\cite{luquette_identification_2019}.
Both estimate the minor allele frequency from nearby germline heterozygous sites. 
However, SCcaller employs a fixed global false positive error rate for the calling of alternative alleles\cite{dong_accurate_2017}, and SCAN-SNV makes use of heuristics for filtering candidate variants\cite{luquette_identification_2019}.
Thus, to the best of our knowledge, there is no statistical model that allows for local variation of bias and errors due to amplification, and for statistically sound false discovery control when calling and genotyping SNVs in single cells.

@@

\begin{table}[tbp]
 \caption{Parameters of the multiple displacement amplification (MDA) amplification bias model as presented in Supplementary Figure S5, panels C and D of \cite{lodato_somatic_2015}. Each of these scales linearly up to a site coverage of 60, with the slopes and intercepts as given here.}
 \label{tab:Lodato}
 \renewcommand{\arraystretch}{1.15}

 \begin{center}
  \begin{tabular}{lrr}
   parameter  & slope        & intersect   \\
   \toprule
   $\alpha$   & -0.000027183 & 0.068567471 \\
   $\beta$    & 0.007454388  & 2.367486659 \\
   \midrule
   $w$        & 0.000548761  & 0.540396786 \\
   $\alpha_1$ & 0.057378844  & 0.669733191 \\
   $\alpha_2$ & 0.003233912  & 0.399261625 \\
   \bottomrule
  \end{tabular}
 \end{center}
\end{table}

@@

\paragraph{Calculating posterior probabilities for events at single cell sites, including a bulk background sample.}
With single cell genotype likelihoods adjusted by our empirical amplification bias model (Equation \ref{eq:BBLodato}) and auxiliary evidence on alternative alleles from a bulk sample (Equation \ref{eq:bulk-prob}), we define mutually exclusive single cell events.
Figure~\ref{fig:prosolo_alt-calling}C gives a simplified illustration of how the combination of likelihood density estimates across allele frequency $\tilde{\theta}_s,\tilde{\theta}_b$ works for calling a heterozygous and a homozygous genotype in the single cell.
However, our model fully defines the two-dimensional space of possible underlying alternative allele frequencies in the two samples as:

@@

We compare ProSolo to the state-of-the-art for SNV calling from single cell sequencing data of multiple displacement amplified (MDA) DNA: MonoVar\cite{zafar_monovar:_2016}, SCAN-SNV\cite{luquette_identification_2019}, SCcaller\cite{dong_accurate_2017} and SCIPhI\cite{singer_single-cell_2018}. We used Snakemake\cite{koster_snakemakescalable_2012} (version {\ttfamily 5.4.0}) to implement the benchmarking workflows.
For detailed information on benchmarking setup and results, see Supplementary Section~\ref{sec:suppl-benchmarking}.
All code used for benchmarking is available at: \url{https://github.com/prosolo/benchmarking_prosolo} (or as preserved by Zenodo at: \url{https://doi.org/10.5281/zenodo.3769115}).

@@

\paragraph{Whole exome sequencing of five human granulocytes with a pedigree ground truth.}
\begin{sloppypar}
For the second benchmarking dataset, blood was taken from a patient with a constitutional mismatch repair-deficiency\cite{hoell_constitutional_2014} after informed consent.
Granulocytes were selected via Magnetic-Activated Cell Sorting (MACS) and Fluorescence Activated Cell Sorting (FACS, Figure~\ref{fig:datasets}B, dataset available as \mbox{EGAD00001005929} in EGA study \mbox{EGAS00001004123}).
Individual cells were isolated using a microfluidics device and subjected to multiple displacement amplification (MDA).
Using a panel of 16 primer pairs covering different genes across chromosomes for quantitative real-time PCR, we selected granulocytes where at least 15 of these loci were properly amplified.
For those cells, we performed whole exome capture, sequencing library preparation and paired-end Illumina sequencing.
From the remaining sorted cell population, we also extracted bulk DNA and submitted it to whole exome capture and paired-end Illumina sequencing without MDA, to generate a bulk background sample for ProSolo and SCcaller.
\end{sloppypar}

@@

\paragraph{Single nucleus exome sequencing of 16 normal and 16 tumor cells from a triple negative breast cancer (TNBC)\cite{wang_clonal_2014}.}
The third dataset is a published dataset from a TNBC patient, with both tumor and matched normal cells available (Figure~\ref{fig:datasets}C, dataset available via the sequence read archive, project id \mbox{PRJNA168068} or accession \mbox{SRA053195}).
For each, 16 single cell nuclei and a bulk sample were whole exome sequenced.
Further, a number of candidate sites for clonal and subclonal tumor mutations was confirmed by targeted duplex deep sequencing in a separate experiment\cite{wang_clonal_2014}.
We analyzed the tumor and normal cells separately, ensuring that normal cells were called with the normal bulk background sample and tumor cells with the tumor bulk background sample.
As the ground truth for the tumor cells, we then used the normal bulk sample augmented with the clonal tumor mutations confirmed by targeted duplex sequencing.
As the ground truth for the normal cells, we used the tumor bulk sample, removing the confirmed clonal and subclonal tumor mutations.
This experimental setup aims for fairness across all competitors. 
Variant calling for the ground truths was performed as for cell line data above.

@@

 \caption{
  Precision-recall plots for alternative allele calls of ProSolo, MonoVar, SCAN-SNV, SCcaller and SCIPhI.
  All panels are strong zoom-ins, focusing on (different) areas of interest.
  Global views of these panels are provided in Figure~\ref{fig:alt-calling_prec-rec_global}.
  \textbf{A} Precision and recall average of two whole genome sequenced single cells IL-11 and IL-12 against their kindred clone IL-1C as ground truth genotypes.
  \textbf{B} Precision and recall average of the five whole exome sequenced single granulocytes against their pedigree-based germline genotype ground truth.
  \textbf{C} Precision and recall average of 16 tumor and 16 normal single cells sequenced at the whole exome level. \newline
  $^\text{\sffamily-B}$ The germline ground truth induces an artificial increase of recall for SCIPhI's sensitive and ProSolo's imputation mode; these modes should thus be disregarded for a fair comparison on the granulocyte dataset in panel B. \newline \footnotesize
  Threshold parameters:
  MonoVar {\ttfamily --t};
  ProSolo {\ttfamily --fdr};
  SCAN-SNV {\ttfamily --fdr};
  SCcaller {\ttfamily -a cutoff};
  SCIPhI {\ttfamily prosolo --fdr}.
  Software modes:
  MonoVar with consensus filtering ({\itshape default}) or without ({\itshape no consensus});
  ProSolo with minimum coverage 1 in single cell ({\itshape default}), or imputing zero coverage sites based on bulk sample ({\itshape imputation});
  SCcaller with recommended settings ({\itshape default}) or with a more {\itshape sensitive} calling;
  SCIPhI with default parameters ({\itshape default}) or all heuristics off ({\itshape sensitive}).
 }
\label{fig:alt-calling_prec-rec}
\end{figure}

@@

At the same time, this germline ground truth caveat also indicates that the recall of SCIPhI's sensitive mode and ProSolo's imputation mode will be an overestimate.
Whenever coverage of a site is missing in a single cell, SCIPhI may impute the genotype with the last common ancestor genotype of the most closely related cells, while ProSolo will impute to the majority genotype in the bulk sample.
Both strategies provide a biologically meaningful imputation that will be more useful than post-hoc modes of imputation.
However, at single cell sites where a somatic mutation has created a true alternative allele, but no coverage is provided, we expect that both methods are most likely to call the homozygous reference germline genotype.
In then comparing this to the germline genotype as the ground truth, these calls will be classified as true negatives even though they really constitute false negatives, thus artificially increasing recall.
While the underestimation of precision equally affects all tools and generally means that benchmarking results are more conservative than with a more accurate (somatic) ground truth, this overestimation of recall in only these modes of two tools does not allow for a fair comparison.
We have thus excluded both SCIPhI's sensitive mode and ProSolo's imputation mode from the discussion of the whole exome dataset with its germline ground truth (but their results are nevertheless displayed in Figures~\ref{fig:alt-calling_prec-rec}B, \ref{fig:alt-calling_prec-rec_global}B, and \ref{fig:F_1-scores_Laehnemann2017} for reference).

For the third dataset of single nucleus whole exome sequencing data of 32 cells from a TNBC patient (16 tumor cells, 16 normal cells), we analysed tumor and normal cells separately (Figure~\ref{fig:alt-calling_prec-rec}C).
On the tumor cells, ProSolo is the only tool to achieve a precision above .99 (at a recall of .319).
On the normal cells, both ProSolo and SCIPhI achieve a precision above .99.
Here, SCIPhI outcompetes ProSolo with a maximum recall of .650 compared to ProSolo's .548.
While ProSolo can achieve a similar recall (.625, see {\ttfamily --fdr} threshold of 0.2 in Figure~\ref{fig:alt-calling_prec-rec_global}C), this comes at reduced precision.
This reflects the inherent uncertainty of the single cell data and showcases a key difference in the approaches of ProSolo and SCIPhI.
Where ProSolo models each cell and each genomic site separately, SCIPhI's model integrates information across all sites in all cells at once.
While this can clearly help recover recall, this can also lead to false positive somatic variant calls (Figure~\ref{fig:somatic-recall}) and has clear implications for model complexity and software runtime.
As all other tools, including ProSolo, are parallelizable over genomic regions and were thus able to process both datasets within days on a multicore machine, we do not report more detailed runtimes for them.
However, it should be noted that SCIPhI took from 1 week (with iterations reduced below software defaults) on the single nucleus whole exome dataset up to 7.5 weeks on the whole genome dataset, running on a single core without any possibility of parallelization.
And where adding breadth of coverage (i.e. more genomic sites) or more cells simply means adding more parallel processes in most tools, both will further increase SCIPhI's wall time.
In addition, adding more cells grows the space of possible tree topologies that SCIPhI explores exponentially\cite{lahnemann_eleven_2020}, which will further increase its runtime.

@@

ProSolo is the first method for SNV calling from MDA single cell sequencing data to comprehensively model both amplification bias and amplification errors in a way that enables both the integration of site-specific differences (Figure \ref{fig:prosolo_alt-calling}), and a computationally efficient---hence practically scalable---evaluation of all relevant data properties.
We achieve this by combining a data-driven model of amplification errors (that incorporates a bulk background sample alongside the single cell sample) with an empirical model of amplification bias which is dependent upon a site's coverage (and is based on a statistical understanding of the MDA process that is mechanistically motivated).
The underlying model calculates posterior probabilities for fine-grained single cell event definitions, whose false discovery rate can be controlled for and that can pass more information about data uncertainties to probabilistic models in downstream analyses; including methods that can then use this reduced, but highly accurate, data representation to integrate information across cells, for example for the phylogenetic reconstruction of cell relationships.
Importantly, the calculation of posterior probabilities scales linearly in the coverage of variant sites, which is the fastest theoretically conceivable, documented by runtimes that have substantial advantages over the state of the art in practice.
Using one whole genome dataset and two whole exome datasets---each with a different type of ground truth (Figure~\ref{fig:datasets})---we demonstrate that these two innovations of ProSolo combined to achieve competing or better accuracy (Figure~\ref{fig:alt-calling_prec-rec}).
Moreover, the model allows to accurately control the false discovery rate of variant calls, another novelty of great practical value, and both the model and its modular implementation are easy to adapt and extend.

### 2021-02-02 manuscript/supplement.tex
last updates to text, e.g. citing the general Zenodo DOI that points to latest pipeline and results data version

\begin{table}[tbp]
 \caption{Parameters of the multiple displacement amplification (MDA) amplification bias model as presented in Supplementary Figure S5, panels C and D of \cite{lodato_somatic_2015}. Each of these scales linearly up to a site coverage of 60, with the slopes and intercepts as given here.}
 \label{tab:Lodato}
 \renewcommand{\arraystretch}{1.15}

 \begin{center}
  \begin{tabular}{lrr}
   parameter  & slope        & intersect   \\
   \toprule
   $\alpha$   & -0.000027183 & 0.068567471 \\
   $\beta$    & 0.007454388  & 2.367486659 \\
   \midrule
   $w$        & 0.000548761  & 0.540396786 \\
   $\alpha_1$ & 0.057378844  & 0.669733191 \\
   $\alpha_2$ & 0.003233912  & 0.399261625 \\
   \bottomrule
  \end{tabular}
 \end{center}
\end{table}

@@

All analysis code for both the main processing pipeline and the manuscript figures is available at: \url{https://github.com/prosolo/benchmarking_prosolo} (or as preserved by Zenodo at: \url{https://doi.org/10.5281/zenodo.3824152}).
An overview and citations of software and resources beyond those mentioned in the methods section are provided in Table~\ref{tab:software-overview}.

\begin{center}
 \begin{table}[tbp]
  \caption{Software and related resources used in this manuscript. All software versions as installed with conda.}
  \label{tab:software-overview}
  \begin{tabular}{llll}
    \toprule
    software / resource & version   & purpose                       & citation \\
    \toprule
    cutadapt            & 1.14      & adapter trimming              & \citep{martin_cutadapt_2011} \\
    \midrule
    trimmomatic         & 0.36      & quality trimming              & \citep{bolger_trimmomatic_2014} \\
    \midrule
    bwa mem             & 0.7.15    & read alignment / mapping      & \citep{li_aligning_2013} \\
    \midrule
    samtools            & 1.4.1     & handle SAM / BAM files        & \citep{li_sequence_2009} \\
    \midrule
    picardtools         & 2.9.2     & mark duplicates               &  \\
    \midrule
    bcftools            & 1.6       & handle VCF / BCF files        & \citep{danecek_variant_2011} \\
    \midrule
    mosdepth            & 0.2.5     & coverage calculations         & \citep{pedersen_mosdepth_2018} \\
    \midrule
    Okabe Ito palette   &           & colorblind-safe               & \citep{okabe_color_2008} \\
    \midrule
    ggthemes            & 0.2.5     & Okabe Ito palette             & \citep{arnold_ggthemes_2019} \\
    \midrule
    Brewer palettes     &           & colorblind-friendly           & \citep{brewer_colorbrewer_2006} \\
    \midrule
    RColorBrewer        & 1.1\_2     & Brewer palettes              & \citep{neuwirth_rcolorbrewer_2014} \\
    \midrule
    tidyverse           & 1.1.1     & ground truth comparisons      & \citep{wickham_welcome_2019} \\
                        & 1.3.0     & results data handling         &  \\
    \midrule
    ggplot2             & 3.2.1     & result plots                  & \citep{wickham_ggplot2_2016} \\
    \midrule
    Inkscape            & 0.92.4    & info-graphics, improved labels & \\
    \midrule
    VGAM                & 1.1\_1     & distributions in R            & \citep{yee_vgam_2008,yee_vector_2015} \\
    \midrule
    UpsetR              &           & set intersection plot \ref{fig:granulocytes-ground-truth} & \citep{conway_upsetr_2017} \\
    \bottomrule
  \end{tabular}
 \end{table}
\end{center}

>>>>

@@

All analysis code for both the main processing pipeline and the manuscript figures is available at: \url{https://github.com/prosolo/benchmarking_prosolo} (or as preserved by Zenodo at: \url{https://doi.org/10.5281/zenodo.3769115}).
An overview and citations of software and resources beyond those mentioned in the methods section are provided in Table~\ref{tab:software-overview}.

\begin{center}
 \begin{table}[tbp]
  \caption{Software and related resources used in this manuscript. All software versions as installed with conda.}
  \label{tab:software-overview}
  \begin{tabular}{llll}
    \toprule
    software / resource & version   & purpose                       & citation \\
    \toprule
    cutadapt            & 1.14      & adapter trimming              & \citep{martin_cutadapt_2011} \\
    \midrule
    trimmomatic         & 0.36      & quality trimming              & \citep{bolger_trimmomatic_2014} \\
    \midrule
    bwa mem             & 0.7.15    & read alignment / mapping      & \citep{li_aligning_2013} \\
    \midrule
    samtools            & 1.4.1     & handle SAM / BAM files        & \citep{li_sequence_2009} \\
    \midrule
    picardtools         & 2.9.2     & mark duplicates               &  \\
    \midrule
    bcftools            & 1.6       & handle VCF / BCF files        & \citep{danecek_variant_2011} \\
                        & 1.10      & some rule-specific environments & \citep{danecek_variant_2011} \\
    \midrule
    mosdepth            & 0.2.5     & coverage calculations         & \citep{pedersen_mosdepth_2018} \\
    \midrule
    Okabe Ito palette   &           & colorblind-safe               & \citep{okabe_color_2008} \\
    \midrule
    ggthemes            & 0.2.5     & Okabe Ito palette             & \citep{arnold_ggthemes_2019} \\
    \midrule
    Brewer palettes     &           & colorblind-friendly           & \citep{brewer_colorbrewer_2006} \\
    \midrule
    RColorBrewer        & 1.1\_2     & Brewer palettes              & \citep{neuwirth_rcolorbrewer_2014} \\
    \midrule
    tidyverse           & 1.1.1     & ground truth comparisons      & \citep{wickham_welcome_2019} \\
                        & 1.3.0     & results data handling         &  \\
    \midrule
    ggplot2             & 3.3.0     & result plots                  & \citep{wickham_ggplot2_2016} \\
    \midrule
    Inkscape            & 0.92.4    & info-graphics, improved labels & \\
    \midrule
    VGAM                & 1.1\_1     & distributions in R            & \citep{yee_vgam_2008,yee_vector_2015} \\
    \midrule
    UpsetR              &           & set intersection plot \ref{fig:granulocytes-ground-truth} & \citep{conway_upsetr_2017} \\
    \bottomrule
  \end{tabular}
 \end{table}
\end{center}
